Benchmarking Python Deep Learning Frameworks\\ for Language Modeling on GPUs

### 2016-12-14 paper/main.tex
DyNet updates

All our frameworks run a \texttt{Python} wrapper\ms{TODO: add footnote about which version of Python}, with a \texttt{C++} backend. Both TensorFlow and DyNet make extensive use of the Eigen library for C++, which provides fast low-level linear algebra for matrix and tensor operations. Theano's backend uses the BLAS linear algebra libraries instead.
\ms{TODO: more spiel? E.g. static computation graph (theano, tf) vs dynamic (Dynet)}
%TensorFlow's computation graph is pure \texttt{Python}, which makes it slower than other frameworks.

@@

Traditionally, neural models are trained until their performance on a \textit{development} set stops improving. Therefore, a full epoch of training (where all the training data is used) is usually followed by a predictive pass through development data.
\section{Benchmarks}
\subsection{Time}
The first benchmark to look at is how long training takes. Table \ref{tab:timing} summarizes various breakdowns. \ms{Summarize results: ie Dynet is the best (so far) (2 sentences perhaps)}.
\begin{table}
\begin{tabular}{c|ccc}
Time					& Dynet 		& TensorFlow 	& Theano \\ \hline
Convergence		& 26h10m 		& 35h25m 			& (TBD)\\
Train epoch 		& 2h20m		& 6h57m 			& 6h35m \\
Test epoch 		& 3m4s			& 7m45s 			& 24m \\
Train batch 		& 1s 				& 5s 					& 1s \\
Test batch 			& $<$1s 		& $<$1s 			& $<$1s \\
\end{tabular}
\caption{\label{tab:timing}Total runtime on various subtasks.}
\end{table}
\subsection{Function-level metrics}
Using NVidia's built-in GPU profiler, \verb!nvprof!, we measured various ways in which the different frameworks utilized the GPU hardware, specifically focusing on the matrix multiply operation implemented in the function \verb!magma_lds128_sgemm_kernel()!. Table \ref{tab:metrics} summarizes. \ms{Summarize results: i.e. they're all in the same order of magnitute, doesn't really tell us why DyNet would be much more efficient.
George TODO: explain the memset thing.}

\begin{table}
\centering
\begin{tabular}{c|rrr}
\textbf{Function} & Dynet &  TensorFlow & Theano \\  
\hline
matrix multiply & 29.64\%  &  68.87\% & 49.90\% \\
\hline
CUDA memset & 40.13\% & 0.00\% & 0.24\% \\
\end{tabular}

@@

\begin{table*}
\centering
\begin{tabular}{c|rrr}
\textbf{Metric} 										& \textbf{Dynet} & \textbf{TensorFlow} & \textbf{Theano} \\ \hline
\texttt{\detokenize{achieved_occupancy}}			&		0.062	&		0.062	&		0.062		\\
\texttt{\detokenize{sm_efficiency}}					&		16.64\%	&		22.58\%	&		16.04\%		\\
\texttt{\detokenize{warp_efficiency}}					&		100.0\%	&		100.0\%	&		99.99\%		\\
\texttt{\detokenize{warp_nonpred_efficiency}}	&		100.0\%	&		99.95\%	&		99.91\%		\\
\texttt{\detokenize{global_hit_rate}}					&		2.88\%	&		0.00\%	&		0.00\%		\\
\texttt{\detokenize{local_hit_rate}}					&		0.00\%	&		0.00\%	&		0.00\%		\\
\texttt{\detokenize{dram_read_throughput}} (GB/s)		&		10.07	&		11.07	 	&		9.97 	\\
\texttt{\detokenize{dram_write_throughput}} (MB/s)	&		357.63&		411.99 &		503.54	\\
\end{tabular}

@@

>>>>

All our frameworks run a \texttt{Python} wrapper\ms{TODO: add footnote about which version of Python}, with a \texttt{C++} backend. Both TensorFlow and DyNet make extensive use of the Eigen library for C++, which provides fast low-level linear algebra for matrix and tensor operations. Theano's backend uses the BLAS linear algebra libraries instead.
\ms{TODO: more spiel? E.g. static computation graph (theano, tf) vs dynamic (DyNet)}
%TensorFlow's computation graph is pure \texttt{Python}, which makes it slower than other frameworks.

@@

Traditionally, neural models are trained until their performance on a \textit{development} set stops improving. Therefore, a full epoch of training (where all the training data is used) is usually followed by a predictive pass through development data.
\section{Benchmarks}
\subsection{Time}
The first benchmark to look at is how long training takes. Table \ref{tab:timing} summarizes various breakdowns. \ms{Summarize results: ie DyNet is the best (so far) (2 sentences perhaps)}.
\begin{table}
\begin{tabular}{c|ccc}
Time					& DyNet 		& TensorFlow 	& Theano \\ \hline
Convergence		& 26h10m 		& 35h25m 			& (TBD)\\
Train epoch 		& 2h20m		& 6h57m 			& 6h35m \\
Test epoch 		& 3m4s			& 7m45s 			& 24m \\
Train batch 		& 1s 				& 5s 					& 1s \\
Test batch 			& $<$1s 		& $<$1s 			& $<$1s \\
\end{tabular}
\caption{\label{tab:timing}Total runtime on various subtasks.}
\end{table}
\subsection{Function-level metrics}
Using NVidia's built-in GPU profiler, \verb!nvprof!, we measured various ways in which the different frameworks utilized the GPU hardware, specifically focusing on the matrix multiply operation implemented in the function \verb!magma_lds128_sgemm_kernel()!. Table \ref{tab:metrics} summarizes. \ms{Summarize results: i.e. they're all in the same order of magnitude, nothing really tells us why DyNet would be much more efficient.} DyNet also spends a large amount of time in the \texttt{CUDA memset} function, unlike the other two frameworks. This may relate to DyNet's memory model: the total amount of memory available to forward and backward passes through the computation graph is specified at runtime, and DyNet may manually manage its memory use in a way TensorFlow and Theano do not.

\begin{table}
\centering
\begin{tabular}{c|rrr}
\textbf{Function} & DyNet &  TensorFlow & Theano \\  
\hline
matrix multiply & 29.64\%  &  68.87\% & 49.90\% \\
\hline
CUDA memset & 40.13\% & 0.00\% & 0.24\% \\
\end{tabular}

@@

\begin{table*}
\centering
\begin{tabular}{c|rrr}
\textbf{Metric} 										& \textbf{DyNet} & \textbf{TensorFlow} & \textbf{Theano} \\ \hline
\texttt{\detokenize{achieved_occupancy}}			&		0.062	&		0.062	&		0.062		\\
\texttt{\detokenize{sm_efficiency}}					&		16.64\%	&		22.58\%	&		16.04\%		\\
\texttt{\detokenize{warp_efficiency}}					&		100.0\%	&		100.0\%	&		99.99\%		\\
\texttt{\detokenize{warp_nonpred_efficiency}}	&		100.0\%	&		99.95\%	&		99.91\%		\\
\texttt{\detokenize{global_hit_rate}}					&		2.88\%	&		0.00\%	&		0.00\%		\\
\texttt{\detokenize{local_hit_rate}}					&		0.00\%	&		0.00\%	&		0.00\%		\\
\texttt{\detokenize{dram_read_throughput}} (GB/s)		&		10.07	&		11.07	 	&		9.97 	\\
\texttt{\detokenize{dram_write_throughput}} (MB/s)	&		357.63&		411.99 &		503.54	\\
\end{tabular}

@@

\paragraph{DyNet}
DyNet includes built-in constructors for various NLP-relevant models, including several varieties recurrent neural nets like Tree LSTMs and encoder-decoders. THis makes it easy for the programmer to build many common kinds of models. DyNet also allows for dynamic computation graph, allowing nodes (representing parameter tensors) to be combined ``on the fly'' to adapt to, for example, data sequences of different lengths. While our neural language model used a static computation graph, DyNet made graph construction easy.

DyNet's primary drawback is the relative lack of documentation and support. The development team and userbase are both fairly small, so there is less information available to help users debug. There is some documentation available online, but it is incomplete.
